---
title: "pilot1"
author: "Samah"
date: "2023-03-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# uncomment below to render to html
# output:
#   bookdown::html_document2:
#     toc: true
#     toc_depth: 4
#     theme: cosmo
#     highlight: tango
    
# uncomment below to render to pdf
output:
  bookdown::pdf_book:
    toc: true
    toc_depth: 4
    highlight: tango
    
bibliography: [references/packages.bib, references/references.bib]
biblio-style: apalike
nocite: '@*'
---



# Results
``` {r, message = F}
# additional libraries
library("knitr")
library("janitor")     
library("broom.mixed") 
library("lme4")
library("emmeans")
library("tidyverse")
library("kableExtra")
library("modelr")
library("broom")
library("nlme")
library("meta")
library("metafor")
library(jtools) # Load jtools
theme_set(theme_classic())
```

```{r, message=FALSE}
# reading the data file
pilot1_data = read_csv("252.csv")
df_shape= filter(pilot1_data, !is.na(d))
# pilot1_data = pilot1_data %>%
#   select(ID, Title, d, d_var, Author)

df_shape_summary = df_shape %>% 
  group_by(ID, Title, Author) %>%
  summarize(mean = mean(d), 
            mean_se = mean(d_var))


  
```

First, Visualizing the data to have an initial idea of how it looks.
First dividing the language group into two groups: the first one is the indo-european group which includes the English and the Spanish languages. The second group includes the rest of the languages: Japanese, Chinese, Tsimane.

``` {r, eval = T, message = F}

df_shape$indoeuropean <- fct_relevel(as.factor(df_shape$language %in% 
                                          c("english","spanish", "german")), 
                                        "TRUE")

# creating a plot that shows the effects sizes colored per language group as well as the polynomial regression curve that fits it. 

ggplot(df_shape,
       aes(x = mean_age_months, y = d, color = language))+
  geom_pointrange(aes(ymin = d - d_var, ymax = d + d_var), 
                  alpha = .5) + 
  geom_smooth(aes(group = indoeuropean, 
                  lty  = indoeuropean), 
              col = "black",
              method = "lm", se = FALSE,
              formula = y ~ poly(x,2)) +
  geom_hline(yintercept = 0, lty = 3) + 
  ylab("Standardized Mean Difference (d)") + 
  xlab("Mean age (months)") +
  scale_color_discrete(name = "Language") +
  scale_linetype_discrete(name = "Indo-European") +
  theme(legend.position = "bottom")

ggsave("first graph.png", width = 7, height = 4)
```
using the meta-analytic function meta-gen which calculates the weights for each effects and confidence interval, pooled effect size, the heterogeneity.
``` {r}
m.gen <- metagen( TE= d, 
                  seTE = d_var, 
                  studlab = ID, 
                  data = df_shape, 
                  sm = "SMD", 
                  fixed = FALSE, 
                  random = TRUE, 
                  method.tau = "REML", 
                  hakn = TRUE, 
                  title = "pilot shape bias meta-analysis"
 
)

summary(m.gen)['TE']
m.gen["TE.fixed"]
m.gen["TE.random"]
m.gen["w.random"]


```

forest plot using the m-gen function object
```{r}
forextobj <- forest.meta(m.gen, 
             sortvar = TE, 
             prediction = TRUE, 
             print.tau2 = FALSE, 
             leftlabs = c("Author", "g", "SE"))

```

forest plot from the rma model:
``` {r, eval = F}

forest(mod) 
forest_data <-tibble(yi = mod$yi,
                     se = sqrt(mod$vi), 
                     slab = mod$slab)

ggplot(forest_data, 
       aes(x = slab, y = yi)) +
  geom_pointrange(aes(ymin = yi - se, ymax = yi + se)) + 
  geom_hline(yintercept = 0, lty = 2) +
  coord_flip() 


# 
# theme_set(theme_bw(base_size=10))
# ata.frame(ES=ROM.ma$b,SE=ROM.ma$se,Type="Summary",Study="Summary"))
# forrest_data$Study2<-factor(forrest_data$Study, levels=rev(levels(forrest_data$Study)) )
# levels(forrest_data$Study2)
# plot1<-ggplot(data=forrest_data,aes(x=Study2,y=ES,ymax=ES+(1.96*SE),ymin=ES-(1.96*SE),size=factor(Type),colour=factor(Type)))+geom_pointrange()
# plot2<-plot1+coord_flip()+geom_hline(aes(x=0), lty=2,size=1)+scale_size_manual(values=c(0.5,1))
# plot3<-plot2+xlab("Study")+ylab("log response ratio")+scale_colour_manual(values=c("grey","black"))
# plot3+theme(legend.position="none")


```


#USING GGPLOT
```{R}
df_shape$mean_age_months_centered36 <- df_shape$mean_age_months - 36
df_shape$log_mean_age_months <- log(df_shape$mean_age_months)

ggplot(df_shape, aes(x = short_cite, y = d, 
                                                             ymin=d-sqrt(d_var)*1.96,
                                                             ymax=d+sqrt(d_var)*1.96)) + 
  geom_pointrange(aes(color=indoeuropean), alpha = .5, position=position_dodge2(width=.5)) +
  coord_flip() +
  geom_hline(yintercept = 0, lty = 2) + 
  aes(x=reorder(short_cite,-d, sum)) +  
  ylab("Standardized Mean Difference (d)") + 
  xlab("Citation")

#png("secondgraph.png")


```


```{r}

 

col.contour = c("gray75", "gray85", "gray95")

funnel(m.gen, 
       comb.random = TRUE,
       xlim = c(-2, 4),
       contour = c(0.9, 0.95, 0.99),
       col.contour = col.contour)

regtest(x = d, vi = d_var,
          data = df_shape)

# Add a legend
legend(x = 3.3, y = 0.1, cex = 0.5,
       legend = c("p < 0.1", "p < 0.05", "p < 0.01"),
       fill = col.contour)


#png("funnel.png")

```

funnel plots using ggplot to account for moderators:
``` {r}
x = summary(m.gen)['TE']
y = summary(m.gen)['seTE']
m.gen["TE.fixed"]
ter = m.gen["TE.random"]


data.gen = data.frame(x,y,ter)
  
ggplot(data = data.gen, mapping = aes(x=TE, y = seTE, color= )) +
  geom_point() +
  geom_vline(xintercept = 0.5401759)


ggplot(data = df_shape, mapping = aes(x=d, y = d_var, color= indoeuropean)) +
  geom_point() +
  geom_vline(xintercept = 0.5401759) +
  scale_y_reverse() +
  geom_smooth(method = "lm")


ggplot(data = df_shape, mapping = aes(x=d, y = d_var, color= language)) +
  geom_point() +
  geom_vline(xintercept = 0.5401759) +
  scale_y_reverse() +
  geom_smooth(method = "lm")


ggplot(data = df_shape, mapping = aes(x=d, y = d_var, color= agegroup)) +
  geom_point() +
  geom_vline(xintercept = 0.5401759) +
  scale_y_reverse()

# +
#   aes(x=d, y =reorder(d_var, -d), color = indoeuropean)


```
using rma.mv instead of m.gen
```{r }
mod <- rma.mv(yi = d, 
                  V = d_var, 
                  random = ~ 1 | ID,
                  slab = short_cite, 
                  data = df_shape)

mod_nested <- rma.mv(yi = d, 
                  V = d_var, 
                  random = ~ 1 | ID/exp_num,
                  slab = short_cite, 
                  data = filter(df_shape, !is.na(exp_num)))

summary(mod)
summary(mod_nested)

```

plotting coefficients from the rmv model:
assuming that those coeffieints correspond to effect sizes
```{r}


```

## Confirmatory analysis

For primary analyses, i will exclude effect sizes from clinical populations and multilingual populations.

I will investigate the hypotheses via multi-level meta-regressions using the metafor package.

In all models, I will include random effects that control for non-independence between effect sizes based on grouping by paper and grouping by experiment.

I will first fit:
Shape bias ~ 1
Shape bias ~ age
shape bias ~ log(age)
shape bias ~ poly(age,2)

### intercept:

``` {r}


# using the meta and metafor packages to analyze meta-analysis effect sizes
mod_intercept <- rma.mv(d ~ 1, 
       V = d_var,
       random = ~1 | as.factor(Title) / 
         as.factor(exp_num), 
       slab = Title, 
       data = filter(df_shape, !is.na(exp_num))) 

summary(mod_intercept)
```
### age
```{r}

mod_age <- rma.mv(d ~ mean_age_months_centered36, 
       V = d_var,
       random = ~1 | as.factor(Title) / 
         as.factor(exp_num), 
       slab = Title, 
       data = filter(df_shape, !is.na(exp_num)))

summary(mod_age)
```
### log age
```{r}

mod_log_age <- rma.mv(d ~ log_mean_age_months, 
       V = d_var,
       random = ~1 | as.factor(Title) / 
         as.factor(exp_num), 
       slab = Title, 
       data = filter(df_shape, !is.na(log_mean_age_months)))

summary(mod_log_age)
```

Let's look at what this means:

```{r}
ggplot(df_shape,
       aes(x = mean_age_months, y = d, color = language))+
  geom_pointrange(aes(ymin = d - d_var, ymax = d + d_var), 
                  alpha = .5) + 
  geom_smooth(aes(group = 1), 
              col = "black",
              method = "lm", se = FALSE,
              formula = y ~ log(x)) +
  geom_hline(yintercept = 0, lty = 3) + 
  ylab("Standardized Mean Difference (d)") + 
  xlab("Mean age (months)") +
  scale_color_discrete(name = "Language") +
  scale_linetype_discrete(name = "Indo-European") +
  theme(legend.position = "bottom") + 
  xlim(0,80)
```

### polynmoial age
```{r}
mod_poly <- rma.mv(d ~ mean_age_months_centered36 + I(mean_age_months_centered36^2), 
       V = d_var,
       random = ~1 | as.factor(Title) / 
         as.factor(exp_num), 
       slab = Title, 
       data = filter(df_shape, !is.na(log_mean_age_months)))

summary(mod_poly)
```

```{r}
m.gen.reg <- metareg(m.gen, ~language)
```

### model comparison and plotting AICc , what is the criteria ? cutoff
```{r}
#anova(mod_log_age, mod_poly, refit = TRUE)
#anova(mod_age, mod_poly) ## the two models are not comparable, not nested

#plot_component(mod, type = "BIC")
summary(mod_intercept)$fit.stats[5,'REML']
AIC_vector <- c(summary(mod_intercept)$fit.stats[5,'REML'],
                summary(mod_age)$fit.stats[5,'REML'], 
                summary(mod_log_age)$fit.stats[5,'REML'],
                summary(mod_poly)$fit.stats[5,'REML'])

# AIC_vector <- c(AICc(mod_intercept), AICc(mod_age), AICc(mod_log_age), AICc(mod_poly))

modelnames <- c("intercept", "age", "logage", "polyage")

data.AIC = data.frame(AIC_vector, modelnames)

ggplot(data = data.AIC, mapping = aes(x = AIC_vector, y = modelnames)) +
  geom_point()

```
setting the contrasts: 

``` {r}

df_shape = df_shape %>% 
  mutate(lang_contrast = factor(language,
                                levels = c("english", "chinese", "german", "japanese", 
                                           "tsimane", "spanish"),
                                labels = c(0,-1,1,1,1,1)),
         lang_contrast = lang_contrast %>% 
           as.character() %>% 
           as.factor() )
contrasts(df_shape$lang_contrast) <- contr.sum(6)
df_shape$lang_contrast
```

### poly age with language
``` {r}
df_shape$lang_factor <- as.factor(df_shape$language)
# contrasts(df_shape$lang_factor) <- contr.treatment(6, base = 2)
df_shape$lang_factor <- fct_relevel(df_shape$lang_factor, "english", after = Inf)
contrasts(df_shape$lang_factor) <- contr.sum(6)*0.5
```
#### orthogonal polynomial:
``` {r}
age_lang <- rma.mv(d ~ poly(mean_age_months_centered36,2) + language , 
                  V = d_var, 
                  random = ~ 1 | ID/exp_num/language,
                  slab = short_cite, 
                  data = filter(df_shape, !is.na(exp_num), !is.na(language)))

age_lang_interact <- rma.mv(d ~ poly(mean_age_months_centered36,2) * language, 
                  V = d_var, 
                  random = ~ 1 | ID/exp_num/language,
                  slab = short_cite, 
                  data = filter(df_shape, !is.na(exp_num), !is.na(language)))

age_lang %>% summary()
age_lang_interact %>% summary()
```

#### the other polynomial:
```{r}

age_lang_npoly <- rma.mv(d ~ mean_age_months_centered36 + I(mean_age_months_centered36^2) + lang_factor, 
                  V = d_var, 
                  random = ~ 1 | ID/exp_num/language,
                  slab = short_cite, 
                  data = filter(df_shape, !is.na(exp_num), !is.na(language)))


age_lang_interact_npoly <- rma.mv(d ~ mean_age_months_centered36 * lang_factor + 
                                    I(mean_age_months_centered36^2) * lang_factor, 
                  V = d_var, 
                  random = ~ 1 | ID/exp_num/lang_factor,
                  slab = short_cite, 
                  data = filter(df_shape, !is.na(exp_num), !is.na(language)))




plot_summs(age_lang_npoly, robust = TRUE)
plot_summs(age_lang_interact_npoly, robust = TRUE)

age_lang_npoly %>% summary()
age_lang_interact_npoly %>% summary()

```

### poly age with syntax

```{r}

age_lang_syntax <- rma.mv(d ~ poly(mean_age_months,2) + language + Informative_syntax, 
                  V = d_var, 
                  random = ~ 1 | ID/exp_num/language,
                  slab = short_cite, 
                  data = filter(df_shape, !is.na(exp_num), !is.na(language)))


rma.mv(d ~ poly(mean_age_months,2)*Informative_syntax, 
       V = d_var,
       random = ~1 | as.factor(Title) + 
         as.factor(exp_num), 
       slab = Title, 
       data = filter(df_shape, !is.na(mean_age_months))) %>% glance()
```

## Exploratory analysis

``` {r}

rma.mv(d ~ poly(mean_age_months,2), 
       V = d_var,
       random = ~1 | as.factor(language)+
         as.factor(Title), 
       slab = Title, 
       data = filter(df_shape, !is.na(mean_age_months), !is.na(language))) %>% glance()
```

# Discussion


# References



