---
title: "MA shapbias analysis"
author: "Samah"
date: "2023-03-14"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
    toc_float: yes
---

# Preliminaries and data reading

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

``` {r, message = F}
# additional libraries
library("knitr")
library("janitor")     
library("broom.mixed") 
library("lme4")
library("emmeans")
library("tidyverse")
library("kableExtra")
library("modelr")
library("broom")
library("nlme")
library(wesanderson)
library("meta")
library("metafor")
library("here")
# library("dmetar")
library(jtools) # Load jtools
theme_set(theme_bw())
```

```{r, message=FALSE}
pilot1_data = read_csv("252.csv")
df_shape <- filter(pilot1_data, !is.na(d))

df_shape_summary <- df_shape %>%
  group_by(language) %>%
  summarize( count = n())

df_shape$englishgrp <- fct_relevel(as.factor(df_shape$language %in% 
                                                 c("english")), 
                                     "TRUE")

df_shape <- df_shape %>% group_by(language) %>%
  mutate(language_grp = case_when(language == "english" ~ "English", 
                                  language %in% c("chinese","japanese", "korean", 
                                                  "vietnamese") ~  "East Asian",
                                  language %in% c("spanish", 
                                                  "german") ~ "Indo-European", 
          TRUE ~ "Tsimane"))


df_shape$mean_age_months_centered36 <- df_shape$mean_age_months - 36
df_shape$log_mean_age_months <- log(df_shape$mean_age_months)

df_shape$indoeuropean <- fct_relevel(as.factor(df_shape$language %in% 
                                                 c("english","spanish", "german")), 
                                     "TRUE")

df_shape_indo <- df_shape %>%
  filter(indoeuropean == TRUE)

df_shape_nonendo <- df_shape %>% 
  filter(indoeuropean == FALSE)

df_shape_solid <- df_shape %>% 
  filter(solid != "substance")
```

# Initial exploration

First, Visualizing the data to have an initial idea of how it looks.

First dividing the language group into two groups: the first one is the indo-european group which includes the English and the Spanish languages. The second group includes the rest of the languages: Japanese, Chinese, Tsimane.

``` {r, eval = T, message = F}

# creating a plot that shows the effects sizes colored per language group as well as the polynomial regression curve that fits it. 


ggplot(df_shape,
       aes(x = mean_age_months, y = d, color = indoeuropean)) + geom_point(aes(ymin = d - d_var, ymax = d + d_var, 
                                                                           alpha = .5, size = part_num)) + 
  geom_smooth(aes(group = indoeuropean, 
                  lty  = indoeuropean), 
              col = "black",
              method = "lm", se = TRUE,
              formula = y ~ poly(x,2)) +
  geom_hline(yintercept = 0, lty = 3) + 
  ylab("Standardized Mean Difference (d)") + 
  xlab("Mean age (months)") +
  scale_color_discrete(name = "Indo-Euro language") +
  scale_linetype_discrete(name = "Indo-European") +
  theme(legend.position = "bottom") + 
  theme_classic(base_size = 8)


ggplot(df_shape,
       aes(x = mean_age_months, y = d, color = language))+
  geom_pointrange(aes(ymin = d - d_var, ymax = d + d_var), 
                  alpha = .5, size = 0.3) + 
  geom_smooth(aes(group = indoeuropean, 
                  lty  = indoeuropean), 
              col = "black",
              method = "lm", se = TRUE,
              formula = y ~ poly(x,2)) +
  geom_hline(yintercept = 0, lty = 3) + 
  ylab("Standardized Mean Difference (d)") + 
  xlab("Mean age (months)") +
  scale_color_discrete(name = "Language") +
  scale_linetype_discrete(name = "Indo-European") +
  theme(legend.position = "bottom") + 
  theme_classic(base_size = 8)


ggplot(df_shape,
       aes(x = mean_age_months, y = d, color = language))+
  geom_pointrange(aes(ymin = d - d_var, ymax = d + d_var), 
                  alpha = .5, size = 0.3) + 
  geom_smooth(aes(group = indoeuropean, 
                  lty  = indoeuropean), 
              col = "black",
              method = "lm", se = TRUE,
              formula = y ~ x) +
  geom_hline(yintercept = 0, lty = 3) + 
  ylab("Standardized Mean Difference (d)") + 
  xlab("Mean age (months)") +
  scale_color_discrete(name = "Language") +
  scale_linetype_discrete(name = "Indo-European") +
  theme(legend.position = "bottom") + 
  theme_classic(base_size = 8)



ggplot(df_shape,
       aes(x = mean_age_months, y = d, color = language_grp))+
  geom_pointrange(aes(ymin = d - d_var, ymax = d + d_var), 
                  alpha = .5, size = 0.3) + 
  geom_smooth(aes(group = language_grp, weight =1/d_var), 
              method = "lm", se = FALSE, 
              formula = y ~ x) +
  geom_hline(yintercept = 0, lty = 3) + 
  ylab("Standardized Mean Difference (d)") + 
  xlab("Mean age (months)") +
  scale_color_discrete(name = "Language") +
  scale_linetype_discrete(name = "Indo-European") +
  theme(legend.position = "bottom") + 
  theme_classic(base_size = 8)



# ggsave("first graph.png", width = 7, height = 4)

barplot(table(df_shape$language), main = "barplot")
length(which(df_shape$language=="english"))

ggplot(df_shape_summary, aes(x = language, y = count)) +
  geom_col(aes(color = language), ) + 
  theme(legend.position = "none")  +
  geom_text(aes(label = count), vjust = -0.2)


```

```{r}


# creating a plot that shows the effects sizes colored per language group as well as the polynomial regression curve that fits it. 

ggplot(df_shape,
       aes(x = mean_age_months, y = d, color = language))+
  geom_pointrange(aes(ymin = d - d_var, ymax = d + d_var), 
                  alpha = .5) + 
  geom_smooth(aes(group = englishgrp, 
                  lty  = englishgrp), 
              col = "black",
              method = "lm", se = FALSE,
              formula = y ~ poly(x,2)) +
  geom_hline(yintercept = 0, lty = 3) + 
  ylab("Standardized Mean Difference (d)") + 
  xlab("Mean age (months)") +
  scale_color_discrete(name = "Language") +
  scale_linetype_discrete(name = "English") +
  theme(legend.position = "bottom") +
  theme_minimal(base_size = 8)


ggplot(df_shape,
       aes(x = mean_age_months, y = d, color = language)) + 
  geom_point(aes(ymin = d - d_var, ymax = d + d_var, 
                  alpha = .5, size = part_num))  +
  geom_smooth(data = filter(df_shape, language == "chinese"), 
              col = "black",
              method = "lm", se = FALSE) + 
  geom_smooth(data = filter(df_shape, language == "english"), 
              col = "black",
              method = "lm", se = FALSE)

```

# Basic meta-analysis

## `metagen` approach

using the meta-analytic function meta-gen which calculates the weights for each effects and confidence interval, pooled effect size, the heterogeneity.

``` {r}
m.gen <- metagen( TE= d, 
                  seTE = d_var, 
                  studlab = ID, 
                  data = df_shape, 
                  sm = "SMD", 
                  fixed = FALSE, 
                  random = TRUE, 
                  method.tau = "REML", 
                  hakn = TRUE, 
                  title = "pilot shape bias meta-analysis"
                  
)

# summary(m.gen)['TE']
# m.gen["TE.fixed"]
# m.gen["TE.random"]
# m.gen["w.random"]
```
### only solid objects: 
``` {r}
m.gen_solid <- metagen( TE= d, 
                  seTE = d_var, 
                  studlab = ID, 
                  data = df_shape_solid, 
                  sm = "SMD", 
                  fixed = FALSE, 
                  random = TRUE, 
                  method.tau = "REML", 
                  hakn = TRUE, 
                  title = "pilot shape bias meta-analysis")
```

forest plot using the m-gen function object

```{r}
forextobj <- forest.meta(m.gen, 
                         sortvar = TE, 
                         prediction = TRUE, 
                         print.tau2 = FALSE, 
                         leftlabs = c("Author", "g", "SE"))
```

forest plot from the rma model

## `metafor` approach

using rma.mv instead of m.gen

```{r }
mod <- rma.mv(yi = d, 
              V = d_var, 
              random = ~ 1 | ID,
              slab = short_cite, 
              data = df_shape)
summary(mod)
```

Nested model. 

```{r}
mod_nested <- rma.mv(yi = d, 
                     V = d_var, 
                     random = ~ 1 | ID/exp_num,
                     slab = short_cite, 
                     data = filter(df_shape, !is.na(exp_num)))

summary(mod_nested)
 

#caluculating the heterogeneity from this website: https://www.metafor-project.org/doku.php/tips:i2_multilevel_multivariate 

W <- diag(1/mod_nested$vi)
X <- model.matrix(mod_nested)
P <- W - W %*% X %*% solve(t(X) %*% W %*% X) %*% t(X) %*% W
100 * sum(mod_nested$sigma2) / (sum(mod_nested$sigma2) + (mod_nested$k-mod_nested$p)/sum(diag(P)))
```

Forest plot from `metafor`.

``` {r}
forest(mod) +
  theme_minimal(base_size = 8)


forest(mod_nested, xlim=c(-16, 4.6), at=log(c(0.05, 0.25, 1, 4)), atransf=exp)
```

Try this using ggplot

```{r, eval=FALSE, include=FALSE}
forest_data <-tibble(yi = mod$yi,
                     se = sqrt(mod$vi), 
                     slab = mod$slab)

ggplot(forest_data, 
       aes(x = slab, y = yi)) +
  geom_pointrange(aes(ymin = yi - se, ymax = yi + se)) + 
  geom_hline(yintercept = 0, lty = 2) +
  coord_flip() 


# 
# theme_set(theme_bw(base_size=10))
# ata.frame(ES=ROM.ma$b,SE=ROM.ma$se,Type="Summary",Study="Summary"))
# forrest_data$Study2<-factor(forrest_data$Study, levels=rev(levels(forrest_data$Study)) )
# levels(forrest_data$Study2)
# plot1<-ggplot(data=forrest_data,aes(x=Study2,y=ES,ymax=ES+(1.96*SE),ymin=ES-(1.96*SE),size=factor(Type),colour=factor(Type)))+geom_pointrange()
# plot2<-plot1+coord_flip()+geom_hline(aes(x=0), lty=2,size=1)+scale_size_manual(values=c(0.5,1))
# plot3<-plot2+xlab("Study")+ylab("log response ratio")+scale_colour_manual(values=c("grey","black"))
# plot3+theme(legend.position="none")


```


## Revised forest plot

```{r, eval=TRUE}

ggplot(df_shape, aes(x = short_cite, y = d, 
                     ymin=d-sqrt(d_var)*1.96,
                     ymax=d+sqrt(d_var)*1.96)) + 
  geom_pointrange( alpha = .5, position=position_dodge2(width=.5)) +
  coord_flip() +
  geom_hline(yintercept = 0, lty = 2) +
  geom_hline(data = m.gen ,yintercept = m.gen$TE.random, color = "red") +
  aes(x=reorder(short_cite,-d, sum)) +  
  ylab("Standardized Mean Difference (d)") + 
  xlab("Citation") +
  theme_minimal(base_size = 8)


 #+geom_text(aes(1.5,m.gen$TE.random,label = round(m.gen$TE.random,2), color = "darkred", size = 0.01 ))

ggplot(df_shape, aes(x = short_cite, y = d, 
                     ymin=d-sqrt(d_var)*1.96,
                     ymax=d+sqrt(d_var)*1.96)) + 
  geom_pointrange(aes(color=indoeuropean), alpha = .5, position=position_dodge2(width=.5)) +
  coord_flip() +
  geom_hline(yintercept = 0, lty = 2) +
  geom_hline(data = m.gen ,yintercept = m.gen$TE.random) +
  aes(x=reorder(short_cite,-d, sum)) +  
  ylab("Standardized Mean Difference (d)") + 
  xlab("Citation") +
  theme_minimal(base_size = 8)

#png("secondgraph.png")
```
## Forest plot with solidity:
```{r}
ggplot(df_shape, aes(x = short_cite, y = d, 
                     ymin=d-sqrt(d_var)*1.96,
                     ymax=d+sqrt(d_var)*1.96)) + 
  geom_pointrange(aes(color=solid), alpha = .5, position=position_dodge2(width=.5)) +
  coord_flip() +
  geom_hline(yintercept = 0, lty = 2) +
  geom_hline(data = m.gen ,yintercept = m.gen$TE.random) +
  aes(x=reorder(short_cite,-d, sum)) +  
  ylab("Standardized Mean Difference (d)") + 
  xlab("Citation") +
  theme_minimal(base_size = 8)


ggplot(df_shape, aes(x = short_cite, y = d, 
                     ymin=d-sqrt(d_var)*1.96,
                     ymax=d+sqrt(d_var)*1.96)) + 
  geom_pointrange(aes(color=solidity), alpha = .5, position=position_dodge2(width=.5)) +
  coord_flip() +
  geom_hline(yintercept = 0, lty = 2) +
  geom_hline(data = m.gen ,yintercept = m.gen$TE.random) +
  aes(x=reorder(short_cite,-d, sum)) +  
  ylab("Standardized Mean Difference (d)") + 
  xlab("Citation") +
  theme_minimal(base_size = 8)
#png("secondgraph.png")
```

```{r}
## only solidity: 

ggplot(df_shape_solid, aes(x = short_cite, y = d, 
                     ymin=d-sqrt(d_var)*1.96,
                     ymax=d+sqrt(d_var)*1.96)) + 
  geom_pointrange(aes(color=solidity), alpha = .5, position=position_dodge2(width=.5)) +
  coord_flip() +
  geom_hline(yintercept = 0, lty = 2) +
  geom_hline(data = m.gen ,yintercept = m.gen_solid$TE.random) +
  aes(x=reorder(short_cite,-d, sum)) +  
  ylab("Standardized Mean Difference (d)") + 
  xlab("Citation") +
  theme_minimal(base_size = 8)


m.gen_solid
m.gen
```

## Publication bias

```{r}

col.contour = c("gray75", "gray85", "gray95")

funnel(m.gen, 
       comb.random = TRUE,
       xlim = c(-3, 6),
       ylim = c(1.05, -0.05),
       contour = c(0.9, 0.95, 0.99),
       col.contour = col.contour)

regtest(x = d, vi = d_var,
        data = df_shape)

# Add a legend
legend(x = 3.3, y = 0.1, cex = 0.5,
       legend = c("p < 0.1", "p < 0.05", "p < 0.01"),
       fill = col.contour)

#png("funnel.png")

funnel(mod_nested, main="Standard Error")
legend(x = 3.3, y = 0.1, cex = 0.5,
       legend = c("p < 0.1", "p < 0.05", "p < 0.01"),
       fill = col.contour)
#funnel(mod_nested, yaxis="seinv", main="Inverse Standard Error")

```

funnel plots using ggplot to account for moderators:

``` {r}
x = summary(m.gen)['TE']
y = summary(m.gen)['seTE']
m.gen["TE.fixed"]
ter = m.gen["TE.random"]


data.gen = data.frame(x,y,ter)

ggplot(data = data.gen, mapping = aes(x=TE, y = seTE, color= )) +
  geom_point() +
  geom_vline(xintercept = 0.4418062) +
  scale_y_reverse()

ggplot(data = df_shape, mapping = aes(x=d, y = d_var, color= indoeuropean)) +
  geom_point() +
  geom_vline(xintercept = 0.5401759) +
  scale_y_reverse()


ggplot(data = df_shape, mapping = aes(x=d, y = d_var, color= language)) +
  geom_point() +
  geom_vline(xintercept = 0.5401759) +
  scale_y_reverse() +
  geom_smooth(method = "lm")

```

## Eggers regresstion test

``` {r}
m.gen$data %>%
  mutate(y = m.gen$TE/m.gen$seTE, x = 1/m.gen$seTE) %>%
  lm(y ~ x, data= .) %>%
  summary()

#eggers regression
ggplot(data = data.gen, mapping = aes(x=1/seTE, y = TE/seTE, color= )) +
  geom_point() +
  geom_smooth(method = "lm")

ggplot(data = df_shape, mapping = aes(x=d_var, y = d/d_var, color= indoeuropean)) +
  geom_point() +
  geom_smooth(method = "lm")

# eggers.test(m.gen)


```

using rma.mv instead of m.gen

```{r }
mod <- rma.mv(yi = d, 
                  V = d_var, 
                  random = ~ 1 | ID,
                  slab = short_cite, 
                  data = df_shape)

mod_nested <- rma.mv(yi = d, 
                  V = d_var, 
                  random = ~ 1 | ID/exp_num,
                  slab = short_cite, 
                  data = filter(df_shape, !is.na(exp_num)))

summary(mod)
summary(mod_nested)

```

plotting coefficients from the rmv model:
assuming that those coefficients correspond to effect sizes

# Confirmatory analysis

For primary analyses, i will exclude effect sizes from clinical populations and multilingual populations.

I will investigate the hypotheses via multi-level meta-regressions using the metafor package.

In all models, I will include random effects that control for non-independence between effect sizes based on grouping by paper and grouping by experiment.

I will first fit:
Shape bias ~ 1
Shape bias ~ age
shape bias ~ log(age)
shape bias ~ poly(age,2)

### intercept:

``` {r}
# using the meta and metafor packages to analyze meta-analysis effect sizes
mod_intercept <- rma.mv(d ~ 1, 
                        V = d_var,
                        random = ~1 | as.factor(Title) / 
                          as.factor(exp_num), 
                        slab = Title, 
                        data = filter(df_shape, !is.na(exp_num))) 


mod_intercept_nonindo <- rma.mv(d ~ 1, 
                        V = d_var,
                        random = ~1 | as.factor(Title) / 
                          as.factor(exp_num), 
                        slab = Title, 
                        data = filter(df_shape_nonendo, !is.na(exp_num))) 

mod_intercept_indo <- rma.mv(d ~ 1, 
                        V = d_var,
                        random = ~1 | as.factor(Title) / 
                          as.factor(exp_num), 
                        slab = Title, 
                        data = filter(df_shape_indo, !is.na(exp_num))) 

summary(mod_intercept_nonindo)
summary(mod_intercept)
```

### age

```{r}

mod_age <- rma.mv(d ~ mean_age_months_centered36, 
                  V = d_var,
                  random = ~1 | as.factor(Title) / 
                    as.factor(exp_num), 
                  slab = Title, 
                  data = filter(df_shape, !is.na(exp_num)))

mod_age_indo <- rma.mv(d ~ mean_age_months_centered36, 
                  V = d_var,
                  random = ~1 | as.factor(Title) / 
                    as.factor(exp_num), 
                  slab = Title, 
                  data = filter(df_shape_indo, !is.na(exp_num)))
summary(mod_age)
summary(mod_age_indo)
```
### log age

```{r}

mod_log_age <- rma.mv(d ~ log_mean_age_months, 
                      V = d_var,
                      random = ~1 | as.factor(Title) / 
                        as.factor(exp_num), 
                      slab = Title, 
                      data = filter(df_shape, !is.na(log_mean_age_months) , !is.na(exp_num)))

mod_log_age_indo <- rma.mv(d ~ log_mean_age_months, 
                      V = d_var,
                      random = ~1 | as.factor(Title) / 
                        as.factor(exp_num), 
                      slab = Title, 
                      data = filter(df_shape_indo, !is.na(log_mean_age_months) , !is.na(exp_num)))

summary(mod_log_age)
summary(mod_log_age_indo)
```

Let's look at what this means:

```{r}
ggplot(df_shape,
       aes(x = mean_age_months, y = d, color = language))+
  geom_pointrange(aes(ymin = d - d_var, ymax = d + d_var), 
                  alpha = .5) + 
  geom_smooth(aes(group = 1), 
              col = "black",
              method = "lm", se = FALSE,
              formula = y ~ log(x)) +
    geom_smooth(aes(group = 1), 
              col = "red",
              method = "lm", se = FALSE,
              formula = y ~ poly(x,2)) + 
  geom_smooth(aes(group = 1), 
              col = "blue",
              method = "lm", se = FALSE,
              formula = y ~ x) +
  geom_hline(yintercept = 0, lty = 3) + 
  ylab("Standardized Mean Difference (d)") + 
  xlab("Mean age (months)") +
  scale_color_discrete(name = "Language") +
  scale_linetype_discrete(name = "Indo-European") +
  theme(legend.position = "bottom") + 
  xlim(0,80)



ggplot(df_shape,
       aes(x = mean_age_months, y = d))+ 
  geom_smooth(aes(group = 1), 
              col = "black",
              method = "lm", se = FALSE,
              formula = y ~ log(x), show.legend = TRUE) +
    geom_smooth(aes(group = 1), 
              col = "red",
              method = "lm", se = FALSE,
              formula = y ~ poly(x,2)) + 
  geom_smooth(aes(group = 1), 
              col = "blue",
              method = "lm", se = FALSE,
              formula = y ~ x) +
  geom_hline(yintercept = 0, lty = 3) + 
  ylab("Standardized Mean Difference (d)") + 
  xlab("Mean age (months)") 
```

### polynomial age

```{r}
mod_poly_nonindo <- rma.mv(d ~ mean_age_months_centered36 + I(mean_age_months_centered36^2), 
                   V = d_var,
                   random = ~1 | as.factor(ID) / 
                     as.factor(exp_num), 
                   slab = Title, 
                   data = filter(df_shape_nonendo, !is.na(log_mean_age_months), !is.na(exp_num)))

mod_poly_indo <- rma.mv(d ~ mean_age_months_centered36 + I(mean_age_months_centered36^2), 
                   V = d_var,
                   random = ~1 | as.factor(ID) / 
                     as.factor(exp_num), 
                   slab = Title, 
                   data = filter(df_shape_indo, !is.na(log_mean_age_months), !is.na(exp_num)))

summary(mod_poly_nonindo)
summary(mod_poly_indo)
```


### model comparison and plotting AICc , what is the criteria ? cutoff

```{r, eval=FALSE, include=FALSE}
#anova(mod_log_age, mod_poly, refit = TRUE)
#anova(mod_age, mod_poly) ## the two models are not comparable, not nested

#plot_component(mod, type = "BIC")
summary(mod_intercept)$fit.stats[5,'REML']
AIC_vector <- c(summary(mod_intercept)$fit.stats[5,'REML'],
                summary(mod_age)$fit.stats[5,'REML'], 
                summary(mod_log_age)$fit.stats[5,'REML'],
                summary(mod_poly)$fit.stats[5,'REML'])

# AIC_vector <- c(AICc(mod_intercept), AICc(mod_age), AICc(mod_log_age), AICc(mod_poly))

models <- c("intercept", "age", "logage", "polyage")

data.AIC = data.frame(AIC_vector, modelnames)

ggplot(data = data.AIC, mapping = aes(x = AIC_vector, y = models, color = models)) +
  geom_point()

# png("modelAIC.png")
# ggsave("modelAIC.png")

```

## Polynomial age models

Let's start with an interaction with `indoeuropean` with standard quadratic terms. This model is very interpretable. 

```{r, include=TRUE}
rma.mv(d ~ mean_age_months_centered36 * indoeuropean + 
         I(mean_age_months_centered36^2) * indoeuropean, 
       V = d_var, 
       random = ~ 1 | ID/exp_num,
       slab = short_cite, 
       data = filter(df_shape, !is.na(exp_num), !is.na(language)))


rma.mv(d ~ mean_age_months_centered36 * indoeuropean , 
       V = d_var, 
       random = ~ 1 | ID/exp_num,
       slab = short_cite, 
       data = filter(df_shape, !is.na(exp_num), !is.na(language)))


rma.mv(d ~ mean_age_months_centered36 * indoeuropean + Informative_syntax , 
       V = d_var, 
       random = ~ 1 | ID/exp_num,
       slab = short_cite, 
       data = filter(df_shape, !is.na(exp_num), !is.na(language)))


```

Next try breaking down by language. Here we can see Spanish is sparse and has a huge interaction term for some reason. Probably just overfit.

```{r, eval=FALSE, include=FALSE}
rma.mv(d ~ mean_age_months_centered36 * language + 
         I(mean_age_months_centered36^2) * language, 
       V = d_var, 
       random = ~ 1 | ID/exp_num,
       slab = short_cite, 
       data = filter(df_shape, !is.na(exp_num), !is.na(language), i))
```
With the orthogonal polynomials, it blows up completely. 

```{r, eval=FALSE, include=FALSE}
rma.mv(d ~ poly(mean_age_months_centered36,2) * language,
       V = d_var, 
       random = ~ 1 | ID/exp_num,
       slab = short_cite, 
       data = filter(df_shape, !is.na(exp_num), !is.na(language)))
```
here is the same model but changing the contrasts:

```{r, eval=FALSE, include=FALSE}
df_shape$lang_factor <- as.factor(df_shape$language)
# contrasts(df_shape$lang_factor) <- contr.treatment(6, base = 2)
df_shape$lang_factor <- fct_relevel(df_shape$lang_factor, "english", after = Inf)
contrasts(df_shape$lang_factor) <- contr.sum(6)*0.5

#1:chinese, 2:german, 3:japanese, 4:spanish, 5:tsimane

rma.mv(d ~ mean_age_months_centered36 * lang_factor + 
         I(mean_age_months_centered36^2) * lang_factor, 
       V = d_var, 
       random = ~ 1 | ID/exp_num,
       slab = short_cite, 
       data = filter(df_shape, !is.na(exp_num), !is.na(lang_factor)))
```

a model that tests for bias ~ age * syntax

``` {r}
rma.mv(d ~ mean_age_months_centered36 * Informative_syntax , 
       V = d_var, 
       random = ~ 1 | ID/exp_num,
       slab = short_cite, 
       data = filter(df_shape, !is.na(exp_num), !is.na(language)))
```
checking the solidity with age : bias ~ exemplar type * age

``` {r}
rma.mv(d ~ mean_age_months_centered36 * solidity , 
       V = d_var, 
       random = ~ 1 | ID/exp_num,
       slab = short_cite, 
       data = filter(df_shape, !is.na(exp_num), !is.na(language)))

```

checking syntax with solidity, not in the preregistration:

``` {r}


```

syntax with count-mass nouns:

``` {r}

rma.mv(d ~ mean_age_months_centered36 + solidity * Informative_syntax, 
       V = d_var, 
       random = ~ 1 | ID/exp_num,
       slab = short_cite, 
       data = filter(df_shape, !is.na(exp_num), !is.na(language)))
```
# alternative grouping
## English , Est Asian , european industrialized , other  :
```{r results = 'asis'}

mod <- rma.mv(d ~ mean_age_months_centered36 * language_grp , 
       V = d_var, 
       random = ~ 1 | ID/exp_num,
       slab = short_cite, 
       data = filter(df_shape, !is.na(exp_num), !is.na(language))) 

rownames(mod$coef) <- c("intercept", "age", "east asian", " indo-european", "Tsimane", "Age * East Asian" , "Age * Indo_European", " Age * Tsiimane ")

# tab_model(mod, pred.labels = c("intercept", "age", "east asian", " indo-european", "Tsimane", "Age * East Asian" , "Age * Indo_European", " Age * Tsiimane "), 
#           dv.labels = "model")


coef(summary(mod)) |>
  kable(digits = 3)

```

```{r  results = 'asis'} 
mod <- rma.mv(d ~ mean_age_months_centered36 * language_grp , 
              V = d_var, random = ~ 1 | ID/exp_num, 
              slab = short_cite, 
              data = filter(df_shape, !is.na(exp_num))) 

summary(mod)

tidy_summary <- tidy(mod)
tidy_summary %>% kable()


coef(summary(mod)) |> 
  papaja::apa_table() 
```
##syntax 
``` {r}
mod_syntax <- rma.mv(d ~ mean_age_months_centered36 * Informative_syntax , 
              V = d_var, random = ~ 1 | ID/exp_num, 
              slab = short_cite, 
              data = filter(df_shape, !is.na(exp_num))) 

summary(mod_syntax)
```
# Discussion


# References



