---
title: "pilot1"
author: "Samah"
date: "2023-03-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# uncomment below to render to html
# output:
#   bookdown::html_document2:
#     toc: true
#     toc_depth: 4
#     theme: cosmo
#     highlight: tango
    
# uncomment below to render to pdf
output:
  bookdown::pdf_book:
    toc: true
    toc_depth: 4
    highlight: tango
    
bibliography: [references/packages.bib, references/references.bib]
biblio-style: apalike
nocite: '@*'
---

```{r setup, echo=FALSE, message=FALSE}
library("knitr") # for knitting things
library("tidyverse") # for all things tidyverse

# these options here change the formatting of how comments are rendered
opts_chunk$set(
  comment = "",
  results = "hold",
  fig.show = "hold")

# set the default ggplot theme 
theme_set(theme_classic())

# include references for used packages
write_bib(.packages(), "references/packages.bib") 
```

# Results
``` {r}
# additional libraries
library("knitr")
library("janitor")     
library("broom.mixed") 
library("lme4")
library("emmeans")
library("tidyverse")
library("kableExtra")
library("modelr")
library("broom")
library("nlme")
library("meta")
library("metafor")
theme_set(theme_classic())
```

```{r}
# reading the data file
pilot1_data = read_csv("252.csv")

# pilot1_data = pilot1_data %>%
#   select(ID, Title, d, d_var, Author)

pilot1_data_summary = pilot1_data %>% 
  group_by(ID, Title, Author) %>%
  summarize(mean = mean(d), 
            mean_se = mean(d_var))
  
```

First, Visualizing the data to have an initial idea of how it looks.
First dividing the language group into two groups: the first one is the indo-european group which includes the English and the Spanish languages. The second group includes the rest of the languages: Japanese, Chinese, Tsimane.

``` {r}

pilot1_data$indoeuropean <- fct_relevel(as.factor(pilot1_data$language %in% 
                                          c("english","spanish", "german")), 
                                        "TRUE")

# creating a plot that shows the effects sizes colored per language group as well as the polynomial regression curve that fits it. 

ggplot(pilot1_data,
       aes(x = mean_age_months, y = d, color = language))+
  geom_pointrange(aes(ymin = d - d_var, ymax = d + d_var), 
                  alpha = .5) + 
  geom_smooth(aes(group = indoeuropean, 
                  lty  = indoeuropean), 
              col = "black",
              method = "lm", se = FALSE,
              formula = y ~ poly(x,2)) +
  geom_hline(yintercept = 0, lty = 3) + 
  ylab("Standardized Mean Difference (d)") + 
  xlab("Mean age (months)") +
  scale_color_discrete(name = "Language") +
  scale_linetype_discrete(name = "Indo-European") +
  theme(legend.position = "bottom")

ggsave("first graph.png", width = 7, height = 4)
```

Starting with all effect sizes without pooling them per paper: In the original dataset, i have multiple effect sizes per paper. So at first i tried to create a forest graph that contains all of the effect sizes without grouping them by paper. 
Note: A forest plot is a graph that provides a graphical display of the observed effect, confidence interval, and usually also the weight of each study. They also display the pooled effect we have calculated in a meta-analysis. 
``` {r}
m.gen <- metagen( TE= d, 
                  seTE = d_var, 
                  studlab = ID, 
                  data = pilot1_data[-9,], 
                  sm = "SMD", 
                  fixed = FALSE, 
                  random = TRUE, 
                  method.tau = "REML", 
                  hakn = TRUE, 
                  title = "pilot shape bias meta-analysis"
 
)

# forest.meta(m.gen, 
#             sortvar = TE, 
#             prediction = TRUE, 
#             print.tau2 = FALSE, 
#             leftlabs = c("Author", "g", "SE"))
```

```{r }
mod <- rma.mv(yi = d, 
                  V = d_var, 
                  random = ~ 1 | ID,
                  slab = short_cite, 
                  data = filter(pilot1_data, 
                                !is.na(d)))
forest(mod) 
forest_data <-tibble(yi = mod$yi,
                     se = sqrt(mod$vi), 
                     slab = mod$slab)

ggplot(forest_data, 
       aes(x = slab, y = yi)) +
  geom_pointrange(aes(ymin = yi - se, ymax = yi + se)) + 
  geom_hline(yintercept = 0, lty = 2) +
  coord_flip() 


# 
# theme_set(theme_bw(base_size=10))
# ata.frame(ES=ROM.ma$b,SE=ROM.ma$se,Type="Summary",Study="Summary"))
# forrest_data$Study2<-factor(forrest_data$Study, levels=rev(levels(forrest_data$Study)) )
# levels(forrest_data$Study2)
# plot1<-ggplot(data=forrest_data,aes(x=Study2,y=ES,ymax=ES+(1.96*SE),ymin=ES-(1.96*SE),size=factor(Type),colour=factor(Type)))+geom_pointrange()
# plot2<-plot1+coord_flip()+geom_hline(aes(x=0), lty=2,size=1)+scale_size_manual(values=c(0.5,1))
# plot3<-plot2+xlab("Study")+ylab("log response ratio")+scale_colour_manual(values=c("grey","black"))
# plot3+theme(legend.position="none")


```


#USING GGPLOT
```{R}
pilot1_data$mean_age_months_centered36 <- pilot1_data$mean_age_months - 36
pilot1_data$log_mean_age_months <- log(pilot1_data$mean_age_months)

ggplot(filter(pilot1_data, !is.na(log_mean_age_months)), aes(x = short_cite, y = d, 
                                                             ymin=d-sqrt(d_var)*1.96,
                                                             ymax=d+sqrt(d_var)*1.96)) + 
  geom_pointrange(aes(color=indoeuropean), alpha = .5, position=position_dodge2(width=.5)) +
  coord_flip() +
  geom_hline(yintercept = 0, lty = 2) + 
  aes(x=reorder(short_cite,-d, sum)) +  
  ylab("Standardized Mean Difference (d)") + 
  xlab("Citation")

#png("secondgraph.png")


```


``` {r}

m.gen_test <- metagen( TE= mean, 
                  seTE = mean_se, 
                  studlab = ID, 
                  data = pilot1_data_summary, 
                  sm = "SMD", 
                  fixed = FALSE, 
                  random = TRUE, 
                  method.tau = "REML", 
                  hakn = TRUE, 
                  title = "pilot shape bias meta-analysis"
)

forest.meta(m.gen_test, 
            sortvar = TE, 
            prediction = TRUE, 
            print.tau2 = FALSE, 
            leftlabs = c("Author", "g", "SE"))

funnel.meta(m.gen,
            xlim = c(-0.5, 2),
            studlab = TRUE)

```

```{r}

 

col.contour = c("gray75", "gray85", "gray95")

funnel.meta(m.gen, xlim = c(-0.5, 2),
            contour = c(0.9, 0.95, 0.99),
            col.contour = col.contour)
png("funnel.png")

```

## Confirmatory analysis

For primary analyses, i will exclude effect sizes from clinical populations and multilingual populations.

I will investigate the hypotheses via multi-level meta-regressions using the metafor package.

In all models, I will include random effects that control for non-independence between effect sizes based on grouping by paper and grouping by experiment.

I will first fit:
Shape bias ~ 1
Shape bias ~ age
shape bias ~ log(age)
shape bias ~ poly(age,2)

``` {r}


# using the meta and metafor packages to analyze meta-analysis effect sizes
mod_intercept <- rma.mv(d ~ 1, 
       V = d_var,
       random = ~1 | as.factor(Title) / 
         as.factor(exp_num), 
       slab = Title, 
       data = filter(pilot1_data, 
                                !is.na(d), !is.na(exp_num))) 

summary(mod_intercept)
```

```{r}

mod_age <- rma.mv(d ~ mean_age_months_centered36, 
       V = d_var,
       random = ~1 | as.factor(Title) / 
         as.factor(exp_num), 
       slab = Title, 
       data = filter(pilot1_data, 
                                !is.na(d), !is.na(exp_num)))

summary(mod_age)
```

```{r}

mod_log_age <- rma.mv(d ~ log_mean_age_months, 
       V = d_var,
       random = ~1 | as.factor(Title) / 
         as.factor(exp_num), 
       slab = Title, 
       data = filter(pilot1_data, !is.na(log_mean_age_months)))

summary(mod_log_age)
```

Let's look at what this means:

```{r}
ggplot(pilot1_data,
       aes(x = mean_age_months, y = d, color = language))+
  geom_pointrange(aes(ymin = d - d_var, ymax = d + d_var), 
                  alpha = .5) + 
  geom_smooth(aes(group = 1), 
              col = "black",
              method = "lm", se = FALSE,
              formula = y ~ log(x)) +
  geom_hline(yintercept = 0, lty = 3) + 
  ylab("Standardized Mean Difference (d)") + 
  xlab("Mean age (months)") +
  scale_color_discrete(name = "Language") +
  scale_linetype_discrete(name = "Indo-European") +
  theme(legend.position = "bottom") + 
  xlim(0,80)
```


```{r}
mod_poly <- rma.mv(d ~ mean_age_months_centered36 + I(mean_age_months_centered36^2), 
       V = d_var,
       random = ~1 | as.factor(Title) / 
         as.factor(exp_num), 
       slab = Title, 
       data = filter(pilot1_data, !is.na(log_mean_age_months)))

summary(mod_poly)
```

```{r}
m.gen.reg <- metareg(m.gen, ~language)
```


```{r}
rma.mv(d ~ poly(mean_age_months,2)*Informative_syntax, 
       V = d_var,
       random = ~1 | as.factor(Title) + 
         as.factor(exp_num), 
       slab = Title, 
       data = pilot1_data[-57,]) %>% glance()
```
## Exploratory analysis

``` {r}

rma.mv(d ~ poly(mean_age_months,2), 
       V = d_var,
       random = ~1 | as.factor(language)+
         as.factor(Title), 
       slab = Title, 
       data = pilot1_data[-57,]) %>% glance()
```

# Discussion


# References



